### 2017-11-13 Draft2/Abstract.tex
[WIP] Feedback

This paper introduces \textbf{growth arrays}, array-like data structures that are designed for appending elements. When the number of items is not known beforehand but is expected to be large, they are more efficient than dynamic arrays by a constant factor. Growth arrays support all operations dynamic arrays do, such as random access, iteration, and insertion or deletion at an index. However, they perform no better, or slightly worse, than dynamic arrays for operations other than appending.

>>>>

This paper introduces \textbf{growth arrays}, array-like data structures that are designed for appending elements. When the number of items is not known beforehand but is expected to be large, growth arrays are more efficient than dynamic arrays by a constant factor. They support all operations dynamic arrays do, such as random access, iteration, and insertion or deletion at an index. For operations other than appending, however, they perform the same as or slightly worse than dynamic arrays.

### 2017-11-13 Draft2/Appending.Dynamic.tex
Make algorithms not float
[WIP] Feedback

\begin{algorithm}
	\begin{algorithmic}[1]
		\Procedure{$\FuncAppend$}{$\VarList,\ \ParamItem$}
			\If{$\VarList.\FieldFull$}
				\State $\VarList.\FuncGrow()$
			\EndIf

@@

% Todo: CITE
It is well-known that appending one item to a dynamic array takes amortized $O(1)$ time. Thus, appending $n$ items takes $O(n)$ time. However, this fact is not very useful for the purpose of comparison--- later, it will be shown that growth arrays also take $O(n)$ time to append $n$ items.

@@

\begin{proof}
	% Todo: Better language here? 'exists' is confusing
	If $\VarCapacitySeq_i$ exists and $i \geq 1$, then clearly $\FuncGrow$ must have been called when the size was $\VarCapacitySeq_{i - 1}$, so $\VarCapacitySeq_{i - 1} \in \VarGrowSeq$. Then $\VarGrowSeq$ contains every term in $\VarCapacitySeq$ except for the last, $\VarCapacitySeq_\VarUseful$, as the corollary states.
\end{proof}

>>>>

\begin{algorithm}[H]
	\begin{algorithmic}[1]
		\Procedure{$\FuncAppend$}{$\VarList,\ \ParamItem$}
			\If{$\VarList.\FieldFull$}
				\State $\VarList.\FuncGrow()$
			\EndIf

@@

It is well-known that appending one item to a dynamic array takes amortized $O(1)$ time. Thus, appending $n$ items takes $O(n)$ time. However, this fact is not very useful for the purpose of comparison--- later, it will be shown that growth arrays also take $O(n)$ time to append $n$ items.

@@

\begin{proof}
	If $\VarCapacitySeq_i$ exists and $i \geq 1$, then clearly $\FuncGrow$ must have been called when the size was $\VarCapacitySeq_{i - 1}$, so $\VarCapacitySeq_{i - 1} \in \VarGrowSeq$. Then $\VarGrowSeq$ contains every term in $\VarCapacitySeq$ except for the last, $\VarCapacitySeq_\VarUseful$, as the corollary states.
\end{proof}

### 2017-11-13 Draft2/Intro.tex
Work
[WIP] Feedback

When dynamic arrays run out of space, they allocate a new buffer, copy the old one's contents into it, then throw the old one away. Growth arrays are less wasteful in this scenario. Instead of discarding the filled buffer, they simply archive it. The new buffer they allocate represents a continuation of the items from the old buffer. For example, if the old buffer contained items $0 - 9$, the new buffer would contain items $10$ and beyond. Because of this technique, growth arrays allocate less memory to store the same number of items, and do not need to copy the contents of old buffers.

Growth arrays have caveats, however. For operations other than appending, they perform no better, or slightly worse, than dynamic arrays. In particular, random access is very costly for growth arrays. They are also not contiguous in memory, which gives them poorer locality than dynamic arrays, and prevents them from being passed to external code that accepts contiguous buffers.

It is worth mentioning that if the size of the data is known in advance, both dynamic and growth arrays are redundant. One could simply allocate a raw array with the known size, and append items to it just as quickly. Thus, growth arrays are only beneficial for cases where the amount of data to be appended is unknown, but is expected to be large.

>>>>

When dynamic arrays run out of space, they allocate a new buffer, copy the old one's contents into it, then throw the old one away. Growth arrays are less wasteful in this scenario. Instead of discarding the filled buffer, they simply archive it. The new buffer they allocate represents a continuation of the items from the old buffer. For example, if the old buffer contained items $0 -- 9$, the new buffer would contain items $10$ and beyond. Because of this technique, growth arrays allocate less memory to store the same number of items, and do not need to copy the contents of old buffers.

Growth arrays have caveats, however. Operations other than appending take the same time or longer than for dynamic arrays. In particular, random access is very costly for growth arrays. They are also not contiguous in memory, which gives them poorer locality than dynamic arrays, and prevents them from being passed to external code that accepts contiguous buffers.

It is worth mentioning that if the size of the data is known in advance, both dynamic and growth arrays are unnecessary. One could simply allocate a raw array with the known size, and append items to it just as quickly. Thus, growth arrays are only beneficial for cases where the amount of data to be appended is unknown, but is expected to be large.
